# Patient characteristics of potential controls

```{r child = "../tables/patchar_all_controls.md"}
```

Table S1. Patient characteristics from all potential controls.

# Dosing and monitoring in subgroups of switchers

```{r child = "../tables/switchers_qoac_subgroups.md"}
```

Table S2. Direct effect of switching in subgroups.  
See also table S6.

# The matching process
##  Propensity scores

```{r child = "../tables/propensity_score_models.md"}
```

Table S3. Coefficients from logistic regression used in the propensity score matching.

## Patient characteristics of matched sets

```{r child = "../tables/patchar_matched.md"}
```

Table S4. Patient characteristics from switchers and selected non-switchers.

# Subgroups
## Patient characteristics

```{r child = "../tables/patchar_subgroups.md"}
```

Table S5. Patient characteristics from switchers and selected non-switchers in subgroups.

## Effect of switching in subgroups

```{r child = "../tables/matched_subgroups.md"}
```

Table S6. Effect of switching from acenocoumarol to phenprocoumon, compared with non-switchers.  
Estimate with 95% confidence interval.  
An asterisk indicates that values were log-transformed in the analyses.
